---
title: "exploratory"
author: "zhuolun"
date: "10/5/2021"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```

## group by name
```{r}
weather_df %>%
  group_by(name, month)
```

## count some things

count month name/observation

```{r}
weather_df %>% 
  group_by(name,month) %>% 
  summarize(n_obs = n())
```

```{r}
weather_df %>% 
  count(name,month,name = "n_obs")
```

**never** use base R's `table`
```{r,eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
## do not use this way 
```

other helpfule counters
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date)
  )
```

## a digression on 2*2 tables

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE    ~ ""
    )
  ) %>% 
  group_by(name,cold) %>% 
  summarize(count = n())
```

this one is easy to use. 
```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ " "
    )
  ) %>%
  janitor::tabyl(name,cold)
```

## more general summarise
let's count ,take means,look at SDs

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax,na.rm = TRUE),
    mean_prec = mean(prcp,  na.rm = TRUE),
    median_tmax = median(tmax,na.rm = TRUE),
    sd_tmax = sd(tmax,na.rm = TRUE)
  )
```

```{r}
weather_df %>%
  group_by(name,month) %>%
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE),
    sd_tmax = sd(tmax, na.rm = TRUE) 
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
  geom_point() +
  geom_line()
```

suppose you want to summarize many columes 

```{r}
weather_df %>%
  group_by(name,month) %>%
  summarize(across(prcp:tmin,mean))
```

reminder sometimes you results are easier to read in another format
```{r}
weather_df %>%
  group_by(name,month) %>%
  summarise(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 1)
```


## `group_by` and `mutate`

```{r}
weather_df %>%
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 

```

what about window functions
```{r}
weather_df %>%
  group_by(name,month) %>%
  mutate(temp_rank = min_rank(desc(tmax)))
```

lagged variables 

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    lagged_tmax = lag(tmax, n = 1),
    tmax_diff = tmax - lagged_tmax
  ) %>% 
  summarize(
    temp_change_sd = sd(tmax_diff, na.rm = TRUE),
    temp_change_max = max(tmax_diff, na.rm = TRUE))
```

## quick note

limitations

what if my summary is a linear model ...
```{r}
weather_df %>% 
  group_by(name) %>% 
  summarise(cor_tmin_tmax = cor(tmin,tamx, use='complete'))

weather_df %>% 
  filter(name = "CentarlPark_NY") %>% 
  lm(tmax - tmin, data =.)
```

