---
title: "ggplot 1"
author: "zhuolun"
date: "10/5/2021"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

## load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

# scatterplot

create my first scatterplot ever

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

new approach same plot
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

save and edit a plot object

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## advanced scatterpolt ..

start with the same one and make it fancy

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

## what about the `aes` placement ?

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) + 
  geom_smooth()
```

let's facet some things 

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax , color = name)) + 
  geom_point(alpha = .2) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

let's combine some elements and try a new plot 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

## Some small notes

how many geoms have to exist ?

you can have whatever geoms you want
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)

```

you can use a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()
```

## use data manlpulation as part of this 
```{r}
weather_df %>%
  filter(name == "CentralPark_NY") %>%
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32 
  ) %>%
  ggplot(aes(x = tmin,y = tmax)) +
  geom_point()
```
```{r}

```

## stacking geoms 


```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```

```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_bin_2d()
```

## Univeriate plots
```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```

histograms are really great

```{r}
weather_df %>%
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin)) +
  geom_histogram() + 
  facet_grid(. ~ name)
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() + 
  facet_grid(. ~ name)
```

let's try a new geometry

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
  
```

what about box plots?
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

trendy plots 
some people like violin plots 
```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmin, fill = name)) +    geom_violin(alpha = .5) +
  stat_summary()
  
```

what about rigdes, most popular plot in 2017 ...

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges(alpha = .8, scake = .8)
```

## save and embedding plots

lets save a scatterplot

```{r,fig.width = 12}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5)

ggsave("./data/weather_plot.pdf",weather_plot, width = 8, height = 5)
```

 what about embeding 
 
```{r,fig,width = 6,fig.height = 3}
weather_plot
```

Embed at different size
```{r,fig.width = 12,fig.height = 6}
weather_plot
```



